Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar

### 2017-05-05 deforestmap.Rmd
Summary

---
title: "Combining global tree cover loss data with national historical forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

1. xxx

2. xxx

@@

In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014. Basically, we used the natural forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007. Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014. We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2010 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes. The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.

>>>>

---
title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

1. The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism. This biodiversity is mainly present in the tropical forests of the island and is highly threatened by anthropogenic deforestation, associated principally to slash-and-burn agriculture and pasture. Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.

2. In this study, we propose a new approach combining recent (2000-2014) global annual tree cover loss data with historical (1953-2000) national forest-cover maps to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar. With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar, for the year 1990 and annually from 2000 to 2014.

3. We estimated that Madagascar has lost 44% of its natural forest on the period 1953-2014 (37% on the period 1973-2014). Natural forest covered 8.9 Mha in 2014 (15% of the national territory) and was divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangroves. Since 2005, deforestation has progressively increased in Madagascar and has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005. Most of the forest is now located at a distance <1.2km of the forest edge while half of it was located far from forest edge (>1km and up to 4km) in 1973. The percentage of most intact forest as then decreased from 68% in 1973 to 50% in 2014.

4. _Conclusion:_ Despite the conservation efforts in Madagascar, deforestation is dramatically increasing at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions. This increase in the deforestation could be attributable to a rapid population growth (close to 3%/yr) and to a high political instability in the country. More effort, including conservation and development programs, should be done in Madagascar to save the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.

@@

In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014. Basically, we used the natural forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007. Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014. We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2005 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes. The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.
